The Accuracy of KNN, SVM, SVDD, neural network and DBSCAN, k-means, NN-chain for Determining the Cancer by Gene-expression Signatures And the Arrhythmia Types

### 2023-06-01 final/paper/main.tex
update paper
update PCA100

The result of gene dataset shows that, with PCA reducing the dimension of the dataset to 32, use KNN for 
classifying the known types, and finally apply DBSCAN on the remaining data, the accuracy of predicting 
can reach about 94\%, which beats the other methods out. In the other way, the result of arrhythmia 
dataset shows that, with raw data, use KNN and k-means, the highest accuracy of predicting can reach about 
53\%.

@@

  On the other hand, KNN performs extraordinarily. In arrhythmia dataset, KNN with k-means in raw data gets 53.1646\% 
  accuracy, which is the highest accuracy comparing to others methods. And in gene expression cancer RNA-Seq dataset, 
  KNN with k-means in PCA data gets 94.8795\%. And KNN with DBSCAN performs well too. This method shows similar performance 
  in these two datasets, which also gets 94.5783\% in gene expression cancer RNA-Seq dataset PCA data.

  For the SVDD + NN with NN-chain, the performance is not as good as KNN, since SVDD cannot fit well with data augmented by 
  interpolation. The hypersphere that defined by SVDD includes all data points in the test dataset, which makes 
  clustering impossible. Therefore, though NN classifying achieves 100\% correctness on the known types, 
  the accuracy of SVDD + NN with NN-chain can only reach 64.45\% in gene expression cancer RNA-Seq dataset.

@@

      Bayes-GA + kmeans           & $0.1342$   & $13.2911 \pm 0$                & $X$        & $X$                 & $0.0294$   & $14.5570 \pm 0$      & $0.0400$   & $31.6456 \pm 0$          & $0.0297$   & $15.8228 \pm 0$      & $0.0399$   & $31.6456 \pm 0$\\
      NN + NN-Chain               & $0.6338$   & $31.6456 \pm 0$                & $0.633 $   & $29.7468 \pm 0$     & $0.9358$   & $31.6456 \pm 0$      & $0.8370$   & $30.3797 \pm 0$          & $1.0540$   & $29.7468 \pm 0$      & $1.0540$   & $29.7468 \pm 0$\\

@@

        Method                  & \multicolumn{2}{c}{Raw}        & \multicolumn{2}{c}{Normalized}     & \multicolumn{2}{c}{PCA-30}                & \multicolumn{2}{c}{PCA-30 Normalized} & \multicolumn{2}{c}{AE-30}         & \multicolumn{2}{c}{AE-30 Normalized} \\ 

@@

      KNN-Brute-Force + DBSCAN  & $14.5101$  & $59.0361 \pm 0$   & $10.3013$  & $\bf{66.5663} \pm 0$  & $0.0224$   & $94.5783 \pm 0$              & $0.0237$   & $78.3133 \pm 0$          & $0.0258$   & $\bf{64.4578} \pm 0$ & $0.0235$   & $59.3373 \pm 0$\\
      SVM (linear) + DBSCAN     & $124.593$  & $61.7470 \pm 0$   & $610.6730$ & $\z{9.9398} \pm 0$    & $0.6466$   & $\z{9.9398} \pm 0$           & $0.1193$   & $10.5422 \pm 0$          & $0.1407$   & $63.2530 \pm 0$      & $0.1122$   & $13.5542 \pm 0$\\
      SVM (polynomial) + DBSCAN & $81.6495$  & $61.7470 \pm 0$   & $86.6406$  & $\z{1.2048} \pm 0$    & $0.2585$   & $64.4578 \pm 0$              & $0.3012$   & $61.1446 \pm 0$          & $0.2013$   & $63.2530 \pm 0$      & $0.2056$   & $50.3012 \pm 0$\\
      SVM (RBF) + DBSCAN        & $X$        & $\text{not converge}$ & $X$    & $\text{not converge}$ & $12.5077$  & $\z{9.9398} \pm 0$           & $3.8838$   & $63.2530 \pm 0$          & $5.0461$   & $40.3614 \pm 0$      & $11.4194$  & $49.6988 \pm 0$\\
      SVM (sigmoid) + DBSCAN    & $X$        & $\text{not converge}$ & $X$    & $\text{not converge}$ & $0.6372$   & $71.6867 \pm 0$              & $0.8222$   & $65.9639 \pm 0$          & $0.6236$   & $44.2771 \pm 0$      & $0.2273$   & $38.2530 \pm 0$\\
      KNN-Brute-Force + kmeans  & $22.754$   & $\bf{74.0964} \pm 0$& $6.36577$& $53.6145 \pm 0$       & $0.0212$   & $\bf{94.8795} \pm 0$         & $0.0214$   & $\bf{84.9398} \pm 0$     & $0.0214$   & $42.1687 \pm 0$      & $0.0220$   & $\bf{65.9639} \pm 0$\\
      SVM (linear) + kmeans     & $136.408$  & $61.7470 \pm 0$   & $664.2860$ & $\z{9.9398} \pm 0$    & $0.5914$   & $\z{9.9398} \pm 0$           & $0.1077$   & $10.5422 \pm 0$          & $0.1274$   & $63.2530 \pm 0$      & $0.1084$   & $13.2530 \pm 0$\\
      SVM (polynomial) + kmeans & $81.4276$  & $61.7470 \pm 0$   & $81.1669$  & $63.8554 \pm 0$       & $0.2003$   & $64.4578 \pm 0$              & $0.2167$   & $25.3012 \pm 0$          & $0.1780$   & $63.2530 \pm 0$      & $0.1930$   & $50.3012 \pm 0$\\
      SVM (RBF) + kmeans        & $1772.32$  & $\z{9.9398} \pm 0$& $1800.6$   & $\z{9.9398} \pm 0$    & $1.3625$   & $\z{9.9398} \pm 0$           & $1.3848$   & $\z{9.9398} \pm 0$       & $11.1729$  & $39.7590 \pm 0$      & $4.7882$   & $28.6145 \pm 0$\\
      SVM (sigmoid) + kmeans    & $190.866$  & $26.5060 \pm 0$   & $81.7487$  & $63.8554 \pm 0$       & $0.8287$   & $65.9639 \pm 0$              & $0.4269$   & $65.9639 \pm 0$          & $0.9908$   & $28.9157 \pm 0$      & $0.2360$   & $39.4578 \pm 0$\\
      Bayes-GA + kmeans         & $2.231$    & $43.6746 \pm 0$   & $X$        & $X$                   & $0.0327$   & $34.0361 \pm 0$              & $0.0364$   & $16.8675 \pm 0$          & $0.0338$   & $27.7946 \pm 0$      & $0.0427$   & $16.8674 \pm 0$\\
      NN + NN-Chain             & $163.359$  & $40.9600 \pm 0$   & $169.031$  & $41.2651 \pm 0$       & $1.1779$   & $35.5421 \pm 0$              & $1.1170$   & $40.96 \pm 0$            & $0.9979$   & $16.4634 \pm 0$      & $1.0220$   & $41.2651 \pm 0$\\

@@

\section{Conclusion}
In this study, we present a framework of algorithm to deal with the dataset containing 
out-of-knowledge data. And with in this algorithm, multiple major machine learning 
techniques are used, e.g., for preprocess, PCA, AE, normalization; for classification, 
KNN, SVM, SVDD, NN, as for clustering, DBSCAN, k-means, NN-chain. Among the various 
combinations, PCA + KNN + k-means performs outstanding in gene expression cancer RNA-Seq 
dataset, and raw data + KNN + k-means performs the best in arrhythmia dataset. Accidently, 
SVM performs bad in these two dataset. The main cause is the defect of the multi-classification 
SVMs design, which make some SVMs hard to separate the target types. However, by customize 
the kernel for each SVM, the problem may ease. In the same time, the increasing number of 
parameters can be fine-tuned by metaheuristic algorithms, e.g., simulated annealing, genetic 
algorithm, etc. We will develope a refinement design on multi-classification SVMs, to enhance 
the accuracy on these datasets. Another issue is SVDD, in 32 dimension PCA data, it can not 
separate the known and unkonwn types well. The reason behind is that in a lower dimension, 
the distance between known and unknown types are so close that SVDD can not separate them. 
If reduce the dataset to a relatively larger dimension, the problem may be solved.

>>>>

The result of gene dataset shows that, with PCA reducing the dimension of the dataset to 32, use KNN for 
classifying the known types, and finally apply DBSCAN on the remaining data, the accuracy of predicting 
can reach about 94\%, which beats the most methods out. And with PCA reducing the dimension to 100, SVDD 
with NN can even have 98\% accuracy. 
On the other way, the result of arrhythmia dataset shows that, with raw data, use KNN and k-means, the 
highest accuracy of predicting can reach about 53\%.

@@

  On the other hand, KNN performs extraordinarily. In arrhythmia dataset, KNN + k-means in raw data gets 53.1646\% 
  accuracy, which is the highest accuracy comparing to others methods. And in gene expression cancer RNA-Seq dataset, 
  KNN + k-means in PCA data gets 94.8795\%. And KNN + DBSCAN performs well too. This method shows similar performance 
  in these two datasets, which also gets 94.5783\% in gene expression cancer RNA-Seq dataset PCA-30 data.

  For the SVDD with NN + NN-chain, the performance is not as good as KNN, since SVDD cannot fit well with data augmented by 
  interpolation. The hypersphere that defined by SVDD includes all data points in the test dataset, which makes 
  clustering impossible. Therefore, though NN classifying achieves 100\% correctness on the known types, 
  the accuracy of SVDD with NN + NN-chain can only reach 64.45\% in gene expression cancer RNA-Seq dataset with PCA-30. 
  However, when in PCA-100, SVDD with NN + k-means and DBSCAN can have 98.4940\% and 97.8916\% respectively.

@@

      Bayes-GA + kmeans           & $0.1342$   & $13.2911 \pm 0$                & $X$        & $\text{not converge}$& $0.0294$  & $14.5570 \pm 0$      & $0.0400$   & $31.6456 \pm 0$          & $0.0297$   & $15.8228 \pm 0$      & $0.0399$   & $31.6456 \pm 0$\\
      SVDD w.t. NN + NN-Chain     & $0.6338$   & $31.6456 \pm 0$                & $0.633 $   & $29.7468 \pm 0$     & $0.9358$   & $31.6456 \pm 0$      & $0.8370$   & $30.3797 \pm 0$          & $1.0540$   & $29.7468 \pm 0$      & $1.0540$   & $29.7468 \pm 0$\\

@@

        Method                  & \multicolumn{2}{c}{Raw}        & \multicolumn{2}{c}{Normalized}     & \multicolumn{2}{c}{PCA-30}                & \multicolumn{2}{c}{PCA-30 Normalized} & \multicolumn{2}{c}{PCA-100} & \multicolumn{2}{c}{PCA-100 Normalized} & \multicolumn{2}{c}{AE-30}         & \multicolumn{2}{c}{AE-30 Normalized} \\ 
        \cmidrule(lr){2-3}\cmidrule(lr){4-5}\cmidrule(lr){6-7}\cmidrule(lr){8-9}\cmidrule(lr){10-11}\cmidrule(lr){12-13}\cmidrule(lr){14-15}\cmidrule(lr){16-17}

@@

      KNN-Brute-Force + DBSCAN  & $14.5101$  & $59.0361 \pm 0$   & $10.3013$  & $\bf{66.5663} \pm 0$  & $0.0224$   & $94.5783 \pm 0$              & $0.0237$   & $78.3133 \pm 0$          & $0.0518$   & $91.2651 \pm 0$& $0.0494$   & $69.8795 \pm 0$           & $0.0258$   & $\bf{64.4578} \pm 0$ & $0.0235$   & $59.3373 \pm 0$\\
      SVM (linear) + DBSCAN     & $124.593$  & $61.7470 \pm 0$   & $610.6730$ & $\z{9.9398} \pm 0$    & $0.6466$   & $\z{9.9398} \pm 0$           & $0.1193$   & $10.5422 \pm 0$          & $4.3810$   & $9.9398 \pm 0$ & $0.5112$   & $9.9398 \pm 0$            & $0.1407$   & $63.2530 \pm 0$      & $0.1122$   & $13.5542 \pm 0$\\
      SVM (polynomial) + DBSCAN & $81.6495$  & $61.7470 \pm 0$   & $86.6406$  & $\z{1.2048} \pm 0$    & $0.2585$   & $64.4578 \pm 0$              & $0.3012$   & $61.1446 \pm 0$          & $0.4586$   & $64.4578 \pm 0$& $0.4612$   & $61.747 \pm 0$            & $0.2013$   & $63.2530 \pm 0$      & $0.2056$   & $50.3012 \pm 0$\\
      SVM (RBF) + DBSCAN        & $X$        & $\text{not converge}$ & $X$    & $\text{not converge}$ & $12.5077$  & $\z{9.9398} \pm 0$           & $3.8838$   & $63.2530 \pm 0$          & $21.3767$  & $9.9398 \pm 0$ & $3.3378$   & $9.6386 \pm 0$            & $5.0461$   & $40.3614 \pm 0$      & $11.4194$  & $49.6988 \pm 0$\\
      SVM (sigmoid) + DBSCAN    & $X$        & $\text{not converge}$ & $X$    & $\text{not converge}$ & $0.6372$   & $71.6867 \pm 0$              & $0.8222$   & $65.9639 \pm 0$          & $0.9179$   & $71.9880 \pm 0$& $1.6702$   & $61.4458 \pm 0$           & $0.6236$   & $44.2771 \pm 0$      & $0.2273$   & $38.2530 \pm 0$\\
      KNN-Brute-Force + kmeans  & $22.754$   & $\bf{74.0964} \pm 0$& $6.36577$& $53.6145 \pm 0$       & $0.0212$   & $\bf{94.8795} \pm 0$         & $0.0214$   & $\bf{84.9398} \pm 0$     & $0.0374$   & $91.8675 \pm 0$& $0.0404$   & $75.6024 \pm 0$           & $0.0214$   & $42.1687 \pm 0$      & $0.0220$   & $\bf{65.9639} \pm 0$\\
      SVM (linear) + kmeans     & $136.408$  & $61.7470 \pm 0$   & $664.2860$ & $\z{9.9398} \pm 0$    & $0.5914$   & $\z{9.9398} \pm 0$           & $0.1077$   & $10.5422 \pm 0$          & $2.3938$   & $9.9398 \pm 0$ & $0.5174$   & $9.9398 \pm 0$            & $0.1274$   & $63.2530 \pm 0$      & $0.1084$   & $13.2530 \pm 0$\\
      SVM (polynomial) + kmeans & $81.4276$  & $61.7470 \pm 0$   & $81.1669$  & $63.8554 \pm 0$       & $0.2003$   & $64.4578 \pm 0$              & $0.2167$   & $25.3012 \pm 0$          & $0.4698$   & $64.4578 \pm 0$& $5.4048$   & $9.9398 \pm 0$            & $0.1780$   & $63.2530 \pm 0$      & $0.1930$   & $50.3012 \pm 0$\\
      SVM (RBF) + kmeans        & $1772.32$  & $\z{9.9398} \pm 0$& $1800.6$   & $\z{9.9398} \pm 0$    & $1.3625$   & $\z{9.9398} \pm 0$           & $1.3848$   & $\z{9.9398} \pm 0$       & $3.1693$   & $9.9398 \pm 0$ & $5.1415$   & $9.9398 \pm 0$            & $11.1729$  & $39.7590 \pm 0$      & $4.7882$   & $28.6145 \pm 0$\\
      SVM (sigmoid) + kmeans    & $190.866$  & $26.5060 \pm 0$   & $81.7487$  & $63.8554 \pm 0$       & $0.8287$   & $65.9639 \pm 0$              & $0.4269$   & $65.9639 \pm 0$          & $1.1448$   & $71.9880 \pm 0$& $1.3232$   & $61.4458 \pm 0$           & $0.9908$   & $28.9157 \pm 0$      & $0.2360$   & $39.4578 \pm 0$\\
      Bayes-GA + kmeans         & $2.231$    & $43.6746 \pm 0$   & $X$        & $\text{not converge}$ & $0.0327$   & $34.0361 \pm 0$              & $0.0364$   & $16.8675 \pm 0$          & $X$        & $X \pm 0$      & $0.0237$   & $X \pm 0$           & $0.0338$   & $27.7946 \pm 0$      & $0.0427$   & $16.8674 \pm 0$\\
      SVDD w.t. NN + NN-Chain   & $163.359$  & $40.9600 \pm 0$   & $169.031$  & $41.2651 \pm 0$       & $1.1779$   & $35.5421 \pm 0$              & $1.1170$   & $40.9600 \pm 0$          & $X$        & $X \pm 0$      & $0.0237$   & $X \pm 0$           & $0.9979$   & $16.4634 \pm 0$      & $1.0220$   & $41.2651 \pm 0$\\
      SVDD w.t. NN + DBSCAN     & $163.359$  & $40.9600 \pm 0$   & $169.031$  & $41.2651 \pm 0$       & $1.1779$   & $35.5421 \pm 0$              & $1.1170$   & $40.9600 \pm 0$          & $X$        & $98.4940 \pm 0$& $0.0237$   & $X \pm 0$           & $0.9979$   & $16.4634 \pm 0$      & $1.0220$   & $41.2651 \pm 0$\\
      SVDD w.t. NN + k-means    & $163.359$  & $40.9600 \pm 0$   & $169.031$  & $41.2651 \pm 0$       & $1.1779$   & $35.5421 \pm 0$              & $1.1170$   & $40.9600 \pm 0$          & $X$        & $97.8916 \pm 0$& $0.0237$   & $X \pm 0$           & $0.9979$   & $16.4634 \pm 0$      & $1.0220$   & $41.2651 \pm 0$\\

@@

\section{Conclusion}
In this study, we present a framework of algorithm to deal with the dataset containing 
out-of-knowledge data. And with in this algorithm, multiple major machine learning 
techniques are used, e.g., for preprocess, PCA, AE, normalization; for classification, 
KNN, SVM, SVDD, NN, as for clustering, DBSCAN, k-means, NN-chain. Among the various 
combinations, PCA + KNN + k-means performs outstanding in gene expression cancer RNA-Seq 
dataset, and raw data + KNN + k-means performs the best in arrhythmia dataset. Accidently, 
SVM performs bad in these two dataset. The main cause is the defect of the multi-classification 
SVMs design, which make some SVMs hard to separate the target types. However, by customize 
the kernel for each SVM, the problem may ease. In the same time, the increasing number of 
parameters can be fine-tuned by metaheuristic algorithms, e.g., simulated annealing, genetic 
algorithm, etc. We will develope a refinement design on multi-classification SVMs, to enhance 
the accuracy on these datasets. Another issue is SVDD, in 32 dimension PCA data, it can not 
separate the known and unkonwn types well, but in 100 dimension PCA data, it can separate konwn 
and unkonwn with 98\% accuracy. The reason behind is because that in a lower dimension, 
the distance between known and unknown types are so close that SVDD can not separate them. 
If reduce the dataset to a relatively larger dimension, the problem can be solved.
